Functional traits complicate the picture of temporal biodiversity change in bird and mammal communities

### 2024-04-22 paper/GEB Submission/manuscript_GEB.tex
Add adjusted p-values

To test for trends within and between different levels of taxa, biome,
realm, and protection status we fit separate models with each of those
factors added as a predictor interacting with time to the original model
structure. We estimated within-level slopes and calculated between-level
contrasts using the \emph{emmeans} package (v1.8.2, 62).
\textcolor{blue}{For some levels of the categorical variables we did not have a sufficient number of studies to estimate a general trend, we therefore refrained from interpreting results for levels where there were fewer than three studies.}
We assessed the impact of time series duration and start year on
study-level trends using linear models with duration and start year as
predictors. All models in our analysis were fit using the \emph{lme4}
(v1.1-30) package in R (v4.2.3) and p-values were calculated by
Satterthwaite's degrees of freedom method using the \emph{lmerTest}
(v3.1-3) package with a significance level of \(\alpha = 0.05\)
(63--65).

@@

We found no significant overall \textcolor{blue}{temporal} trend in
species richness or functional diversity metrics
\textcolor{blue}{including functional richness, evenness, or divergence}
(observed or \textcolor{blue}{corrected}) (Fig
\ref{fig:timeseriesPlot}). We did find a significant overall decrease in
Jaccard similarity, indicating accumulating changes in species
composition. Non-significant overall \textcolor{blue}{temporal} trends
indicate that although some studies experience increasing or decreasing
trends, the average trend across studies
\textcolor{blue}{was not significantly different from zero} (Table
\ref{tab:resultsTab}). \textcolor{blue}{Within-group} trends for
different taxa, biomes, realms, or protection statuses were also
non-significant for richness and functional diversity metrics, with the
exception of a significantly increasing trend for functional evenness of
marine studies, and a significantly decreasing functional richness slope
for mammal studies. However, trends were not exhibited for the
\textcolor{blue}{corrected} metric indicating that differences
\textcolor{blue}{in functional diversity metrics} were largely due to
changes in species richness. The general trends for \emph{CWM} models
were similarly not significant, with a significant positive trend for
only percentage of fish in diet composition.

@@

We did find significant differences between taxa, realms, and biomes for
Jaccard similarity and some of the \emph{CWM's}. For example, while
Jaccard similarity was decreasing in the general trend and there were
significant within group slopes for Mammals, Birds, Terrestrial,
Temperate, and Tropical communities, there was no significant slope for
the marine realm, indicating that the general trend is mostly driven by
turnover in terrestrial communities. We also found a significant
decrease in Jaccard similarity in unprotected areas only, with no trend
for protected areas. We found significant dietary shifts across
communities, with a significant general trend of increasing fish
consumption, which was also reflected in increasing fish consumption
trends for Bird studies and Temperate studies. We found a significantly
increasing trend in nectar consumption for Temperate studies but a
significantly decreasing tend for Tropical studies. We found
significantly decreasing trend for vertebrate consumption in Marine
studies and Tropical studies.

At the study level, 16 of 50 studies exhibited a significant trend in
species richness and 11 exhibited significant turnover. For uncorrected
functional diversity metrics, 25 of 50 studies exhibited a trend in a
least one metric, and 15 of 50 studies exhibited a significant trend for
at least one \textcolor{blue}{corrected} metric (Table
\ref{tab:trendTab}). In general, there were more significant trends for
uncorrected metrics, with some disappearing after correction, indicating
that those trends were likely due to changes in the number of species.
Hypothesis testing for study-level trends is likely affected by multiple
testing issues and some trends identified as significant are therefore
potentially spurious. Rather than interpreting changes in specific
studies, we present these results as a general picture of the kinds of
trends experienced by communities.

@@

\caption{\label{tab:trendTab}Number of studies that experienced a significant trend in each calculated metric out of 50 total studies.}
\centering
\begin{tabular}[t]{ccccccccc}
\toprule
 & S & Jaccard Similarity & FRic & FEve & FDiv & SES FRic & SES FEve & SES FDiv\\
\midrule
+ & 7 & 0 & 7 & 5 & 3 & 4 & 4 & 3\\
- & 9 & 11 & 8 & 6 & 5 & 2 & 6 & 2\\
\bottomrule
\end{tabular}
\end{table}

>>>>

To test for trends within and between different levels of taxa, biome,
realm, and protection status we fit separate models with each of those
factors added as a predictor interacting with time to the original model
structure. We estimated within-level slopes and calculated between-level
contrasts using the \emph{emmeans} package (v1.8.2, 62).
\textcolor{blue}{For some levels of the categorical variables we did not have a sufficient number of studies to estimate a general trend, we therefore only interpret results for levels with more than three studies.}
We assessed the impact of time series duration and start year on
study-level trends using linear models with duration and start year as
predictors. All models in our analysis were fit using the \emph{lme4}
(v1.1-30) package in R (v4.2.3) and p-values were calculated by
Satterthwaite's degrees of freedom method using the \emph{lmerTest}
(v3.1-3) package with a significance level of \(\alpha = 0.05\)
(63--65).

@@

We found no significant overall \textcolor{blue}{temporal} trend in
species richness or functional diversity metrics
\textcolor{blue}{including functional richness, evenness, or divergence}
(observed or \textcolor{blue}{corrected}) (Fig
\ref{fig:timeseriesPlot}). We did find a significant overall decrease in
Jaccard similarity, indicating accumulating changes in species
composition. Non-significant overall \textcolor{blue}{temporal} trends
indicate that although some studies experience increasing or decreasing
trends, the average trend across studies
\textcolor{blue}{was not significantly different from zero} (Table
\ref{tab:resultsTab}). \textcolor{blue}{Within-group} trends for
different taxa, biomes, realms, or protection statuses were also
non-significant for richness and functional diversity metrics. The
general trends for \emph{CWM} models were similarly not significant,
with a significant positive trend for only percentage of fish in diet
composition.

@@

We did find significant differences between taxa, realms, and biomes for
Jaccard similarity and some of the \emph{CWM's}. For example, while
Jaccard similarity was decreasing in the general trend and there were
significant within group slopes for birds, terrestrial, and temperate
studies, there was no significant slope for marine, mammal, or tropical
studies. We also found a significant decrease in Jaccard similarity in
unprotected areas only, with no trend for protected areas. We found
significant dietary shifts across communities, with a significant
general trend of increasing fish consumption, which was also reflected
in increasing fish consumption trends for bird studies and temperate
studies. We found significantly decreasing trend for vertebrate
consumption in Marine studies and Tropical studies. The trend for marine
vertebrate consumption was significantly more negative than for
terrestrial studies and the temperate trend was significantly more
negative than tropical studies.

At the study level, 12 of 50 studies exhibited a significant trend in
species richness and 11 exhibited significant turnover. For uncorrected
functional diversity metrics, 11 of 50 studies exhibited a trend in a
least one metric, and 7 of 50 studies exhibited a significant trend for
at least one \textcolor{blue}{corrected} metric (Table
\ref{tab:trendTab}). In general, there were more significant trends for
uncorrected metrics, with some disappearing after correction, indicating
that those trends were likely due to changes in the number of species.

@@

\caption{\label{tab:trendTab}Number of studies that experienced a significant trend in each calculated metric out of 50 total studies.}
\centering
\begin{tabular}[t]{ccccccccc}
\toprule
 & S & Jaccard Similarity & FRic & FEve & FDiv & SES FRic & SES FEve & SES FDiv\\
\midrule
+ & 2 & 0 & 5 & 3 & 1 & 4 & 0 & 1\\
- & 9 & 12 & 2 & 0 & 3 & 1 & 1 & 2\\
\bottomrule
\end{tabular}
\end{table}
